On the uniqueness and stability of dictionaries \\ for sparse representation of noisy signals

### 2019-05-10 manuscripts/IEEE/final_submission.tex
edit start

% TODO: 
% cite Fritz l0 paper in introduction for seemingly characteristic structure? 
% Make L_H a definition..? Change def of L to be 1-norm so we don't need these sqrt k's?
% C_1 depends on J in Thm 1
% also in proof of (7) x need not be *any* 2k-sparse vector...can instead restrict to some H...then 2k is restricted to half being on H.

@@

\section{Introduction}\label{Intro}
%\IEEEPARstart{L}{earning} sparse representations of data has become an important step in modern solutions to problems in signal processing and pattern analysis. 
% \cite{sato1975method}
\IEEEPARstart{A}{}common modern approach to pattern analysis in signal processing is to view each of $N$ observed $n$-dimensional signal samples as a (noisy) linear combination of at most $k$ elementary waveforms drawn from some unknown ``dictionary" of size $m \ll N$ (see \cite{Zhang15} for a comprehensive review). 
Optimizing dictionaries subject to this and related sparsity constraints has revealed seemingly characteristic sparse structure in several signal classes of current interest (e.g., in vision \cite{wang2015sparse}).

@@

Motivated by these discoveries and more recent work relating compressed sensing \cite{eldar2012compressed} to a theory of information transmission through random wiring bottlenecks in the brain \cite{Isely10}, we address when optimal dictionaries for sparse representation are indeed identifiable from data. Answers to this question may also have implications in practice wherever an appeal is made to latent sparse structure of data (e.g., forgery detection \cite{hughes2010, olshausen2010applied}; brain recordings \cite{jung2001imaging, agarwal2014spatially, lee2016sparse}; and gene expression \cite{wu2016stability}).

@@

Here, we prove very generally that uniqueness and stability is a typical property of sparse dictionary learning. More specifically,  we show that dictionaries injective on a sparse domain are identifiable from \mbox{$N = m(k-1){m \choose k} + m$} noisy sparse linear combinations of their columns up to an error that is linear in the noise (Thm.~\ref{DeterministicUniquenessTheorem}). In fact, provided $n \geq \min(2k,m)$, in almost all cases the problem is well-posed, as per Hadamard \cite{Hadamard1902}, given a sufficient amount of data (Thm.~\ref{robustPolythm} and Cor.~\ref{ProbabilisticCor}).

>>>>

% TODO: 
% Make L_H a definition..? Change def of L to be 1-norm so we don't need these sqrt k's?
% C_1 depends on J in Thm 1
% also in proof of (7) x need not be *any* 2k-sparse vector...can instead restrict to some H...then 2k is restricted to half being on H.

@@

\section{Introduction}\label{Intro}
\IEEEPARstart{A}{}common modern approach to pattern analysis in signal processing is to view each of $N$ observed $n$-dimensional signal samples as a (noisy) linear combination of at most $k$ elementary waveforms drawn from some unknown ``dictionary" of size $m \ll N$ (see \cite{Zhang15} for a comprehensive review). 
Optimizing dictionaries subject to this and related sparsity constraints has revealed seemingly characteristic sparse structure in several signal classes of current interest (e.g., in vision \cite{wang2015sparse}).

@@

Motivated by these discoveries and more recent work relating compressed sensing \cite{eldar2012compressed} to a theory of information transmission through random wiring bottlenecks in the brain \cite{Isely10}, we address when dictionaries for sparse representation are indeed identifiable from data. Answers to this question may also have implications in practice wherever an appeal is made to latent sparse structure of data (e.g., forgery detection \cite{hughes2010, olshausen2010applied}; brain recordings \cite{jung2001imaging, agarwal2014spatially, lee2016sparse}; and gene expression \cite{wu2016stability}).

@@

Here, we prove very generally that uniqueness and stability is a typical property of sparse dictionary learning. More specifically,  we show that dictionaries injective on a sparse domain are identifiable from \mbox{$N = m(k-1){m \choose k} + m$} noisy k-sparse linear combinations of their $m$ columns up to an error that is linear in the noise (Thm.~\ref{DeterministicUniquenessTheorem}). In fact, provided $n \geq \min(2k,m)$, in almost all cases the problem is well-posed, as per Hadamard \cite{Hadamard1902}, given a sufficient amount of data (Thm.~\ref{robustPolythm} and Cor.~\ref{ProbabilisticCor}).
